Live Functional Programming with Typed Holes

### 2018-04-30 hazelnut-dynamics.tex
suppress ACM information

\PassOptionsToPackage{svgnames,dvipsnames,svgnames}{xcolor}
%% For double-blind review submission
\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=false,printccs=false,printacmref=false}
%% For single-blind review submission
%\documentclass[acmlarge,review]{acmart}\settopmatter{printfolios=true}
%% For final camera-ready submission
%\documentclass[acmlarge]{acmart}\settopmatter{}

@@

%% Listings
\makeatletter\if@ACM@journal\makeatother
%% Journal information (used by PACMPL format)
%% Supplied to authors by publisher for camera-ready submission
\acmJournal{PACMPL}
\acmVolume{1}
\acmNumber{1}
\acmArticle{1}
\acmYear{2018}
\acmMonth{3}
\acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}
\else\makeatother
%% Conference information (used by SIGPLAN proceedings format)
%% Supplied to authors by publisher for camera-ready submission
\acmConference[PL'18]{ACM SIGPLAN Conference on Programming Languages}{January 01--03, 2017}{New York, NY, USA}
\acmYear{2018}
\acmISBN{978-x-xxxx-xxxx-x/YY/MM}
\acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}
\fi

@@

@@

%% \maketitle
%% Note: \maketitle command must come after title commands, author
%% commands, abstract environment, Computing Classification System
%% environment and commands, and keywords command.
\maketitle

>>>>

\PassOptionsToPackage{svgnames,dvipsnames,svgnames}{xcolor}
%% For double-blind review submission
\documentclass[acmsmall,review]{acmart}
\settopmatter{printfolios=false,printccs=false,printacmref=false}
\setcopyright{none}
\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference information in first column

%% For single-blind review submission
%\documentclass[acmlarge,review]{acmart}\settopmatter{printfolios=true}
%% For final camera-ready submission
%\documentclass[acmlarge]{acmart}\settopmatter{}

@@

\makeatletter\if@ACM@journal\makeatother
%% Journal information (used by PACMPL format)
%% Supplied to authors by publisher for camera-ready submission
\acmJournal{PACMPL}
\acmVolume{1}
\acmNumber{1}
\acmArticle{1}
\acmYear{2018}
\acmMonth{3}
\acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}
\else\makeatother
%% Conference information (used by SIGPLAN proceedings format)
%% Supplied to authors by publisher for camera-ready submission
% \acmConference[]{ACM SIGPLAN Conference on Programming Languages}{January 01--03, 2017}{New York, NY, USA}

\acmYear{2018}
\acmISBN{978-x-xxxx-xxxx-x/YY/MM}
\acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}
\fi

@@

\fancyfoot{} % suppresses the footer (also need \thispagestyle{empty} after \maketitle below)

@@

%% \maketitle
%% Note: \maketitle command must come after title commands, author
%% commands, abstract environment, Computing Classification System
%% environment and commands, and keywords command.
\maketitle
\thispagestyle{empty} % suppresses the footer

### 2018-04-30 related-work.tex
suppress ACM information

Our approach is reminiscent of the workflow that debuggers make available using breakpoints \cite{fitzgerald2008debugging,DBLP:journals/jfp/TolmachA95} and a variety of logging and tracing capabilities. Debuggers do not directly support incomplete programs, so the programmer first needs to insert suitable dummy exceptions as discussed in Sec.~\ref{sec:intro}. Beyond that, there are two main distinctions. First, evaluation does not stop at each breakpoint and so it is straightforward to explore the \emph{space} of values that a variable takes. Second, breakpoints, logs and tracing tools convey the values of a variable at a position in the evaluation trace. Hole closures, on the other hand, convey information from a syntactic position in the \emph{result} of evaluation. The result is, by nature, a simpler object.%  These approaches are fundamentally complementary in that standard debugging facilities are useful even when there are no holes in the program whereas our focus has been on incomplete programs.
Some debuggers support ``edit-and-continue'', e.g. Visual Studio \cite{VSEditAndContinue} and Flutter \cite{flutter}, based on the dynamic software update (DSU) capabilities of the underlying run-time system \cite{DBLP:journals/toplas/StoyleHBSN07,DBLP:conf/vstte/HaydenMHFF12}. These features do not come with any guarantee that rerunning the program is  guaranteed to produce the same result. The ``fill-and-resume'' feature described in Sec.~\ref{sec:resumption} is a special case that comes with a stronger formal guarantee.

>>>>

Our approach is reminiscent of the workflow that debuggers make available using breakpoints \cite{fitzgerald2008debugging,DBLP:journals/jfp/TolmachA95} and a variety of logging and tracing capabilities. Debuggers do not directly support incomplete programs, so the programmer first needs to insert suitable dummy exceptions as discussed in Sec.~\ref{sec:intro}. Beyond that, there are two main distinctions. First, evaluation does not stop at each breakpoint and so it is straightforward to explore the \emph{space} of values that a variable takes. Second, breakpoints, logs and tracing tools convey the values of a variable at a position in the evaluation trace. Hole closures, on the other hand, convey information from a syntactic position in the \emph{result} of evaluation. The result is, by nature, a simpler object.%  These approaches are fundamentally complementary in that standard debugging facilities are useful even when there are no holes in the program whereas our focus has been on incomplete programs.
Some debuggers support ``edit-and-continue'', e.g. Visual Studio \cite{VSEditAndContinue} and Flutter \cite{flutter}, based on the dynamic software update (DSU) capabilities of the underlying run-time system \cite{DBLP:journals/toplas/StoyleHBSN07,DBLP:conf/vstte/HaydenMHFF12,DBLP:journals/toplas/HicksN05}. These features do not come with any guarantee that rerunning the program will produce the same result. The ``fill-and-resume'' feature described in Sec.~\ref{sec:resumption} is a special case that comes with this stronger guarantee.
